US President Prediction

### 2024-10-22 paper/paper.qmd
Update the paper.qmd

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

@@

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

@@

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

@@

# Model

@@

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

@@

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

@@

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

@@

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

@@

@@

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

>>>>

---
title: "US President Prediction"
subtitle: "Predict the results of 2024 US President Election"
author: 
  - Yun Chu
  - Felix Li
  - Wen Han Zhao 
thanks: "Code and data are available at: [https://github.com/younazhao/US-President-Prediction/tree/main](https://github.com/younazhao/US-President-Prediction/tree/main)."
date: 22 October 2024
date-format: long
abstract: "In this study, we aim to study the 2024 U.S. Presidential Election with the methodology of  polls of polls across various states. With the insightful analyses on polls provided by Redfiled & Wilton Strategies, the conducted survey on the population highlights various key point for the prediction of US president. Using Generalized Linear Models (GLMs), this report analyzes the accuracy of prediction in president votes through several key variables from pollster. The model incorporates variables such as pollscore, methodology, transparency_score and hypothetical. The prediction finding highlights the significance of candicate trustworthiness among the voter decision, providing insights for swing voters. This analysis offers valuable information into the potential electoral outcomes driving the 2024 US president election. "
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
library(tidyverse)
library(knitr)
library(kableExtra)
library(readr)
library(dplyr)
```

@@

We use the statistical programming language R [@citeR].... Our data.... Following @tellingstories, we consider...

@@

@@

# Model {#set-model}

@@

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix @sec-model-details].

@@

Define $y_i$ as the percentage of support for a candicate in a poll.

${y_i} = \beta_0 + \beta_1 pollscore + \beta_2 methodology + \beta_3transparency score + \beta_4 hypothetical + \epsilon$

pollscore: A numeric value representing the score or reliability of the pollster in question (e.g., -1.1). "The error and bias we can attribute to a pollster. Negative numbers are better. Stands for "Predictive Optimization of Latent skill Level in Surveys, Considering Overall Record, Empirically.

methodology: The method used to conduct the poll (e.g., Online Panel).

transparency_score: A score reflecting the pollster’s transparency about their methodology (e.g., 9.0). "A grade for how transparent a pollster is, calculated based on how much information it discloses about its polls and weighted by 
by recency. The highest Transparency Score is 10."

hypothetical: Indicates whether the poll is about a hypothetical match-up.

We run the model in R [@citeR].

@@

@@

# Library setup
#install.packages("glm")
library(glmnet)
library(ggplot2)
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-model1
#| tbl-cap: "model 1"
#| warning: false
# Get the analysis data
analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

# Model
glm_model1 <- glm(pct ~ pollscore + methodology + transparency_score + hypothetical, data = analysis_data, family = gaussian())

# Get the summary
model_summary <- summary(glm_model1)

# Convert the summary coefficients to a data frame
summary_table <- as.data.frame(model_summary$coefficients)

# Create a table with a caption and control the size
kable(summary_table, caption = "Model 1 Summary Table", size = "small")
```

## Lasso Regularization {#sec-model-lassoregularization}
Lasso Regularization is performed to see if the selected variable in our model would best predict the results of polls percentage of a candidate. If the variable do not align with our actual numbers, we should consider other variables or a interaction variable.

```{r}
#| echo: false
#| eval: true
#| label: tbl-lassoregularization
#| tbl-cap: "lasso regularization"
#| warning: false
# Prepare the predictors
x <- model.matrix(pct ~ pollscore + methodology + transparency_score + hypothetical, data = analysis_data)[, -1]

# Prepare the response variable
y <- analysis_data$pct

# Lasso regression (alpha = 1)
lasso_model <- glmnet(x, y, alpha = 1, family = "gaussian")

# Cross-validation for Lasso
cv_lasso <- cv.glmnet(x, y, alpha = 1, family = "gaussian")

plot(cv_lasso)

# Get the best lambda
best_lambda_lasso <- cv_lasso$lambda.min

# Refit Lasso with the optimal lambda
lasso_final_model <- glmnet(x, y, alpha = 1, lambda = best_lambda_lasso, family = "gaussian")

# View Lasso coefficients
coef.lasso <- coef(lasso_final_model)

# Predict using Lasso
pred_lasso <- predict(lasso_final_model, s = best_lambda_lasso, newx = x)

```

Computing the Mean Square Error would give us a sense of our lasso model prediction. 
```{r}
#| echo: false
#| eval: true
#| label: tbl-msr
#| tbl-cap: "Mean Squared Error"
#| warning: false
# Calculate Mean Squared Error
mse <- mean((y - pred_lasso)^2)
cat("Mean Squared Error:", mse, "\n")

# Calculate R-squared
sst <- sum((y - mean(y))^2)
sse <- sum((y - pred_lasso)^2)
r_squared <- 1 - (sse / sst)
cat("R-squared:", r_squared, "\n")

# Create a data frame for plotting
results_df <- data.frame(
  Metric = c("SST", "SSE", "R-squared"),
  Value = c(sst, sse, r_squared)
)

kable(results_df, caption = "Mean Squared Error Table")
```

From this graph, we can see that the cluster points do not perfectly align on the line of the best fit so we should still investigate in other variables. 
```{r}
#| echo: false
#| eval: true
#| label: tbl-actualvspredicted
#| tbl-cap: "actual vs predicted"
#| warning: false

# Compare predicted vs actual values
comparison <- data.frame(Actual = y, Predicted = as.vector(pred_lasso))

kable(head(comparison), caption = "Actual vs Predicted")

# Plot actual vs predicted values
ggplot(comparison, aes(x = Actual, y = Predicted)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  labs(title = "Actual vs Predicted Values",
       x = "Actual Values",
       y = "Predicted Values") +
  theme_minimal()
```

@@

## Weakness & Strength of the methodology

In terms of strengths, Redfield & Wilton has a great reputation for producing reliable polling data. They have employ a mix of online and telephone survey, which add in various resources of collecting their data. This approach can help reduce bias. In addition, Redfield & Wilton often target swing states, which makes their polls results important to the US president election.

The weakness of their methodology would incorporate a certain potential bias based on their political leaning of their clients or media. This could cause a certain neutrality in their dataset.

Overall, Redfield & Wilton has been considered as a competent, reputable and reliable pollster with his variaty and methodology on polls. Even though the pollster contained a potential bias, it has been a reliable resource in prediction of US president election.

@@
